Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities

### 2020-10-09 writing/transect_paper.Rmd
specify samples sizes explicitly to clean up low-sample values on accumulation curves, and move OTU curve to supplement.
make read count comparison flow better in results

We compared the unique ASVs and OTUs shared between datasets from different sequencing strategies, and the number of reads represented by these ASVs and OTUs in each strategy.
The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Tables\ \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu)).
The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).
When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% of reads, respectively (Figure\ \@ref(fig:venn)b).
In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs, were found to be shared with other sequencing technologies upon OTU clustering.
ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` of belonged to a unique OTU after clustering.
In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure\ \@ref(fig:venn)c), and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after clustering the ITS2 regions into OTUs (Figure\ \@ref(fig:venn)d).

Figure\ \@ref(fig:read-compare) shows the correspondence between the read count for different ASVs (\@ref(fig:read-compare)a) and OTUs(\@ref(fig:read-compare)b) in the different technologies, where shared ASVs/OTUs are plotted as circles, and unshared OTUs are plotted as lines along the axes.
In all cases, the read counts for shared ASVs and OTUs were correlated, with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")`.
Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
(`r compR2("pb_483", "SH-2257", "ASV")` to `r compR2("pb_483", "SH-2257", "OTU")`,
`r compR2("pb_483", "is_057", "ASV")` to `r compR2("pb_483", "is_057", "OTU")`, and
`r compR2("SH-2257", "is_057", "ASV")` to `r compR2("SH-2257", "is_057", "OTU")`,
for PacBio vs. Illumina, PacBio vs. Ion Torrent, and Illumina vs. Ion Torrent, respectively), but not between the long amplicon library and short amplicon library
(`r compR2("pb_500", "pb_483", "ASV")` to `r compR2("pb_500", "pb_483", "OTU")`,
`r compR2("pb_500", "SH-2257", "ASV")` to `r compR2("pb_500", "SH-2257", "OTU")`, and
`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs. PacBio, Illumina, and Ion Torrent short amplicon reads, respectively).

@@

```{r accum-OTU, fig.cap="OTU accumulation curves"}
readd(accum_plot_OTU)
```

>>>>

We compared the unique ASVs and OTUs shared between datasets from different sequencing strategies, and the number of reads represented by these ASVs and OTUs in each strategy.
The majority of abundant ASVs and OTUs were captured by all sequencing strategies used (Tables\ \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu)).
The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).
When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))`% of reads, respectively (Figure\ \@ref(fig:venn)b).
In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.
ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` belonged to a unique OTU after clustering.
In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure\ \@ref(fig:venn)c), and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after clustering the ITS2 regions into OTUs (Figure\ \@ref(fig:venn)d).

Read counts for shared ASVs and OTUs were highly correlated between strategies, with a minimum $R^2$ value of `r compR2("pb_500", "is_057", "ASV")` (Figure\ \@ref(fig:read-compare)).
Correlations between read counts for the three technologies using the short amplicon library were increased by OTU clustering
(`r compR2("pb_483", "SH-2257", "ASV")` to `r compR2("pb_483", "SH-2257", "OTU")`,
`r compR2("pb_483", "is_057", "ASV")` to `r compR2("pb_483", "is_057", "OTU")`, and
`r compR2("SH-2257", "is_057", "ASV")` to `r compR2("SH-2257", "is_057", "OTU")`,
for PacBio vs. Illumina, PacBio vs. Ion Torrent, and Illumina vs. Ion Torrent, respectively), but not between the long amplicon library and short amplicon library
(`r compR2("pb_500", "pb_483", "ASV")` to `r compR2("pb_500", "pb_483", "OTU")`,
`r compR2("pb_500", "SH-2257", "ASV")` to `r compR2("pb_500", "SH-2257", "OTU")`, and
`r compR2("pb_500", "is_057", "ASV")` to `r compR2("pb_500", "is_057", "OTU")`, for PacBio long amplicon reads vs. PacBio, Illumina, and Ion Torrent short amplicon reads, respectively).

<!-- ADD HERE results about the alpha diversity comparison -->

@@
